## Checks

```{r}

source("utils.R")
library(tidyverse)

ppts.raw <- read.csv("data/vipv_participant.csv")
binary <- read.csv("data/vipv_binary.csv")

# Preprocess participants
ppts.all <- preprocess.ppts(
  ppts.raw, "pilot")


pv.all <- binary %>%
  filter(participant_id %in% ppts.all$id,
         task == "PV") %>%
  mutate(accuracy = ifelse(is_correct == "True", 1, 0))

pv.critical.all <- pv.all %>% filter(item_type == "critical")

wm.all <- binary %>%
  filter(participant_id %in% ppts.all$id,
         task == "WM") %>%
  mutate(accuracy = ifelse(is_correct == "True", 1, 0))

wm.critical.all <- wm.all %>% filter(item_type == "critical")

wm.dt <- wm.critical.all %>%
  select(participant_id, trial_id, block_id, condition, version,
         reaction_time, accuracy) %>%
  rename(
    wm.modality = condition,
    wm.load = version,
    wm.reaction_time = reaction_time,
    wm.accuracy = accuracy
  )

vipv <- merge(pv.critical.all %>% mutate(block_id = as.integer(block_id)), wm.dt, by=c("trial_id", "block_id", "participant_id"), all = F) %>%
  rename (
    pv.modality = condition,
    pv.accuracy = accuracy,
    pv.reaction_time = reaction_time,
  ) %>%
  mutate(
    pv.modality = factor(pv.modality, levels=c("vi", "au")),
    wm.modality = factor(wm.modality, levels=c("vi", "au")),
    acc.dt = pv.accuracy + wm.accuracy,
    pv.rt.log = log(pv.reaction_time),
    pv.rt.log.c = scale(pv.rt.log),
    wm.rt.log = log(wm.reaction_time),
    wm.rt.log.c = scale(wm.rt.log),
  ) %>%
  arrange(trial_id)


```

```{r}

binary %>%
  filter(participant_id == "17", item_type=="critical",
         trial_id == "1_7"
         ) 

```

```{r}

vipv %>% filter(participant_id == 17, trial_id=="1_7")

```

```{r}
# CF list 1 csv

vipv %>% filter(participant_id == 17, item_type=="critical") %>%
  select(item, pv.modality, wm.modality, wm.load)

```

# 2-way interactions

```{r}

m.vipv.pv.acc.2.base <- glmer(
    pv.accuracy ~ 1 + pv.modality + wm.modality + wm.load + 
              (1 | participant_id) + (1 | item_id),
  data = vipv, family="binomial", control=optimxglmerControl)

m.vipv.pv.acc.2 <- glmer(
  pv.accuracy ~ 1 + pv.modality + wm.modality + wm.load + 
    pv.modality:wm.modality +
    (1 | participant_id) + (1 | item_id),
  data = vipv, family="binomial", control=optimxglmerControl)

summary(m.vipv.pv.acc.2)
anova(m.vipv.pv.acc.2.base, m.vipv.pv.acc.2)

```


# Property Verification

## Item sampling

Items appear to be sampled in a balanced way.

```{r}

pv.critical %>%
  ggplot(aes(x = reorder(item_id, accuracy), fill=condition)) + 
  geom_histogram(stat="count") + 
  theme_minimal() + 
  theme(
    axis.text.x = element_blank()
  )

```

# Working Memory

## Item sampling

Items appear to be sampled in a balanced way.

```{r}

wm.critical %>%
  ggplot(aes(x = item_id, fill=condition)) + 
  geom_histogram(stat="count") + 
  theme_minimal() + 
  theme(
    axis.text.x = element_blank()
  )

```

## Accuracy

Retained participants performed better on social vs visual domain items.

```{r}

wm.critical %>%
  ggplot(aes(x = condition, y = accuracy)) + 
  facet_grid(rows=vars(version)) +
  stat_summary(fun.data = "mean_cl_boot", geom="errorbar", width=0.2) +
  stat_summary(fun = "mean", geom="point", size=3) + 
  theme_minimal()
  

```


```{r}

m.wm.condition <- lmer(
  accuracy ~ 1 + condition + (1 + condition | participant_id) + (1 | item_id),
  data = wm.critical
)

summary(m.wm.condition)

```

```{r}

m.wm.version <- lmer(
  accuracy ~ 1 + version + (1 + version | participant_id) + (1 | item_id),
  data = wm.critical
)

summary(m.wm.version)

```


# IRQ

#### IRQ

```{r}

irq.attention.answers <- data.frame(
  item_id = c(37, 38, 39),
  correct_response = c(1,5,1))

irq.all <- irq.all %>%
  mutate(item_type = recode(item_type, catch = "attention"))

ppts.all <- exclude.ppts.attention(ppts.all, irq.all, irq.attention.answers, "irq", cutoff.irq.attention)

```

`r round(nrow(ppts.all %>% filter(irq.attention.accuracy>0.66)) / nrow(ppts.all), 2) * 100`% of ppts got > 66% IRQ attention checks correct.

```{r}

ppts.all %>%
  ggplot(aes(x = irq.attention.accuracy)) + 
  geom_histogram() + 
  theme_minimal() + 
  labs(
    x = "IRQ Attention Accuracy",
    y = "No. Participants"
  )

```

Participants who failed IRQ attention check.

```{r}

ppts.all %>%
  filter(irq.attention.accuracy < 0.66) %>%
  select(participant_id, irq.attention.accuracy)

```

IRQ Accuracy by question

```{r}

irq.all %>%
  filter(item_type == "attention") %>%
  merge(irq.attention.answers) %>%
  mutate(accuracy = ifelse(response == correct_response, 1, 0)) %>%
  group_by(item_id) %>%
  summarize(accuracy = mean(accuracy), .groups="drop")

```